
```{r}
library('ggplot2')
```

```{r}
df = read.csv('master.csv')
sapply(df, function(x) mean(is.na(df)))
```

Next ANOVA test is telling us that there is significant different group between the generation groups.

But gives no information about which one has the highest suicide rate.

```{r}
gen.aov <- aov(suicides_no~generation, data=df)
summary(gen.aov)
```

```{r}
pairwise.t.test(df$suicides_no, as.vector(df$generation))
```

there are significant different from each other.

To find the highest one we test data with t test.

According to the plot we saw Boomers generation is the highest one in term of suicide rate.

ANOVA+Pairwise T-test gives us the result that Boomers generation is the highest in term of suicide rate worldwide.

```{r}
df %>% 
  ggplot(aes(y = generation,
             x = suicides_no,
             fill = generation) ) +
  geom_boxplot(alpha = .50) +
  coord_cartesian(xlim = c(0,500))
```

Next ANOVA test is telling us that there is significant difference between sexes.

But gives no information about which one has the highest suicide rate.

```{r}
gen.aov <- aov(suicides_no~sex, data=df)
summary(gen.aov)
```

We run t-test to compare 2 sexes and see if males are more prone to committing suicide then females

```{r}
pairwise.t.test(df$suicides_no, as.vector(df$sex))
```

Also obvious comparison for 2 samples is plotting.

Results are proven. We see that males are more prone to committing suicides.

```{r}
df %>% 
  ggplot(aes(y = sex,
             x = suicides_no,
             fill = sex) ) +
  geom_boxplot(alpha = .50) +
  coord_cartesian(xlim = c(0,500))
```
